Toronto Homicides and Shootings (To be Changed)

### 2024-09-29 paper/paper.qmd
Updated references and paper

The remainder of this paper is structured as follows. @sec-data....

@@

### Neighbourhood Crime Rates Dataset
The "cleaned_crime_data" dataset was retrieved from the "Neighbourhood Crime Rates" dataset and the last time the dataset was updated on the Open Data Portal was on January 11, 2024. This dataset comprised of crime data by neighbourhood that included the number of cases of assault, robbery, homicide, shootings, auto theft, break and enter, and theft over. The original dataset was published by the Toronto Police Services collected from 2014 to 2023, inclusive and is updated annually. For the cleaned dataset, we decided to filter for homicides and shootings and focus on the cases from 2018 to 2023, inclusive. The 2023 population projection in the dataset was provided by Environics Analytics where they counted only the residents in the neighbourhood [@citeneighbourhoodcrimerates].

Each observation in the dataset represent one of Toronto's 158 neighbourhoods and its shootings and homicides from the years 2018 to 2023, inclusive that was recorded by the Toronto Police Services. There are 1896 unique observations in total in the dataset.

@@

# Add up total number of cases each year across all neighbourhoods
yearly_crime <- 
  cleaned_crime_data |>
  unique() |>
  # Calculate the total sum of cases across all neighbourhoods 
  mutate(
    num_cases_total = sum(num_of_cases)
  ) |>
  # Group data by year, type of crime, number of total cases, and number of cases
  select(year, homicide_or_shooting, num_cases_total, num_of_cases) |>
  group_by(year, homicide_or_shooting) |>
  unique() |>
  # Create new columns for number of cases per year for homicides and shootings across ALL neighbourhood. Also calculate the proportion as well. 
  mutate (
    yearly_num_cases_total = sum(num_of_cases),
    yearly_num_cases_total_prop = round(yearly_num_cases_total / num_cases_total, 2)
  ) |>
  tibble() |>
  select(year, homicide_or_shooting, yearly_num_cases_total, yearly_num_cases_total_prop) |>
  unique() |>
  # Arrange based on the total number of cases per year 
  arrange(desc(yearly_num_cases_total)) |>
  ungroup()
```
## Number of Homicide and Shooting Cases in Toronto from 2018 to 2023 in Toronto

@@

# Discussion {#sec-discussion}

@@

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

@@

Weaknesses and next steps should also be included.

@@

# Appendix {-}

>>>>

The introduction is self-contained and tells a reader everything they need to know including: 1) broader context to motivate (done); 2) some detail about what the paper is about (done); 3) a clear gap that needs to be filled (done); 4) what was done; 5) what was found; 6) why it is important; 7) the structure of the paper (done). A reader should be able to read only the introduction and know what was done, why, and what was found. Likely 3 or 4 paragraphs, or 10 per cent of total.

As a response to the increase in gun violence in 2024 by 74% from 2023, the Toronto Police Services announced on June 27, 2024 at a press conference that they will increase their presence in areas impacted by high levels of gun violence [@citeincreasepolicetoronto]. This concern was expressed because in a CBC article published in December 2023, they mentioned that there was a decrease in gun violence from 2019 to 2023 across Toronto [@citedropgunviolence]. Deputy Chief Rob Johnson in the same conference also raised concerns on the increasing number of youth being perpetrators of gun violence [@citeyouthgunviolence]. This raises the following questions: Which Toronto neighbourhoods are most impacted by shootings and homicides from 2018 to 2023? Where are police stations located relative to the shootings and homicides across Toronto? According to Statistics Canada, the most frequently used method to commit homicides is through shooting and so in our analysis we chose to analyze homicides alongside shootings [@citestatcan].

In this paper, we analyzed data provided by Toronto Police Services and City Planning to investigate patterns where shootings and homicides take place and where police stations are located relative to where the crimes occur. Currently there is news coverage about increasing police visibility in areas with high gun violence but it is not known if the location of police stations in Toronto does impact the number of shootings and homicides that have taken place.

In our findings,.....

Our findings can also inform police divisions of neighbourhoods to be more present at based on how many cases has occurred in a neighbourhood.

For the rest of the paper, the data section [@sec-data] will describe the datasets used and they were retrieved, our variables of interest used in our graphs and tables, and briefly about any variables that were constructed and the data cleaning process. In the results section [@sec-results], it will describe and show graphs and tables created based on our datasets and in the discussion section [@sec-discussion], it will provide an overview of what we did in our results, explain the meaning of our results and the implications it has on the real world, and provide potential areas of improvements of the paper and suggestions of future works.

@@

### Neighbourhood Crime Rates Dataset
The "cleaned_crime_data" dataset was retrieved from the "Neighbourhood Crime Rates" dataset and the last time the dataset was updated on the Open Data Portal was on January 11, 2024. This dataset comprised of crime data by neighbourhood that included the number of cases of assault, robbery, homicide, shootings, auto theft, break and enter, and theft over. The original dataset was published by the Toronto Police Services collected from 2014 to 2023, inclusive and is updated annually. For the cleaned dataset, we decided to filter for homicides and shootings and focus on the cases from 2018 to 2023, inclusive. The 2023 population projection in the dataset was provided by Environics Analytics where they only included the number of residents in the neighbourhood [@citeneighbourhoodcrimerates].

Each observation in the dataset represent one of Toronto's 158 neighbourhoods and its shootings and homicides from the years 2018 to 2023, inclusive that was recorded by the Toronto Police Services as crimes occur in Toronto. The original dataset is updated annually by the Toronto Police Services in January of the following year. There are 1896 unique observations in total in the dataset.

@@

# Add up total number of cases each year across all neighbourhoods
yearly_crime <- 
  cleaned_crime_data |>
  unique() |>
  # Calculate the total sum of cases across all neighbourhoods 
  mutate(
    num_cases_total = sum(num_of_cases)
  ) |>
  # Group data by year, type of crime, number of total cases, and number of cases
  select(year, homicide_or_shooting, num_cases_total, num_of_cases) |>
  group_by(year, homicide_or_shooting) |>
  unique() |>
  # Create new columns for number of cases per year for homicides and shootings across ALL neighbourhood. Also calculate the proportion as well. 
  mutate (
    yearly_num_cases_total = sum(num_of_cases),
    yearly_num_cases_total_prop = round(yearly_num_cases_total / num_cases_total, 2)
  ) |>
  tibble() |>
  select(year, homicide_or_shooting, yearly_num_cases_total, yearly_num_cases_total_prop) |>
  unique() |>
  # Arrange based on the total number of cases per year 
  arrange(desc(yearly_num_cases_total)) |>
  ungroup()
```
## Number of Homicides and Shootings in Toronto from 2018 to 2023 in Toronto

@@

# Discussion {#sec-discussion}
In @sec-results, we focused on cases from 2018 to 2023 and looked at the number of homicides and shootings yearly and the six Toronto neighbourhoods that had the highest number of homicides and shootings using tables and graphs. Finally, we mapped where police facilities are located in Toronto alongside the number of homicides and shootings for each neighbourhood.

@@

Neighbourhoods with the highest number of shootings and homicides 
- Glenfield-Jane Heights
- Mount Olive-Silverstone-Jamestown	
- Black Creek	
- York University Heights	
- Yorkdale-Glen Park
- West Humber-Clairville

https://www.npr.org/sections/money/2021/04/20/988769793/when-you-add-more-police-to-a-city-what-happens

https://www.mcgilldaily.com/2023/02/more-policing-does-not-address-the-rise-of-violent-crime-in-canada/#:~:text=The%20NBER%20research%20found%20that,deeply%20harmful%20to%20marginalized%20communities

@@

We faced several limitations while conducting our analysis. We do not have information about when police facility were opened.

@@

# Appendix {#sec-appendix}
